---
title: "Generate UMAP and color in UMAP by metadata"
authors: "Nayanga Thirimanne"
---
 

```{r}
library(ggplot2)
install.packages("umap")
library(umap)
library(tibble)
library(readxl)
library(dplyr)

setwd("~/Output_VSTcounts/")
```

ggplot2 specifications
```{r}
plot_title_size = 35
legend_pt_size = 25
axis_text_size = 23
axis_title_size = 27
legend_text_size = 15
spacing = 1
chosen_margin = c(0.5,1,0.5,1) #top,right,bottom,left

theme_legend <- theme_void() +
  theme(plot.title = element_text(hjust=0, vjust=0, lineheight=.8, face="bold", size=plot_title_size),
        plot.margin = unit(chosen_margin,"cm"),
        legend.text = element_text(size=legend_text_size, face = "bold"),
        legend.key.height = unit(spacing,"cm"),
        legend.position = "right",
        legend.justification = "left",
        legend.title = element_blank())

theme_nolegend <- theme_void() +
  theme(plot.title = element_text(hjust=0, vjust=0, lineheight=.8, face="bold", size=plot_title_size),
        plot.margin = unit(chosen_margin,"cm"),
        legend.text = element_text(size=legend_text_size, face = "bold"),
        legend.key.height = unit(spacing,"cm"),
        legend.position = "none",
        legend.justification = "left",
        legend.title = element_blank())

theme_grid_nolegend <- theme_bw()+
  theme(plot.title = element_text(hjust=0, vjust=0, lineheight=.8, face="bold", size=plot_title_size),
        plot.margin = unit(chosen_margin,"cm"),
        legend.text = element_text(size=legend_text_size, face = "bold"),
        legend.key.height = unit(spacing,"cm"),
        legend.position = "none",
        legend.justification = "left",
        legend.title = element_blank(), 
        axis.text =  element_text(size = axis_text_size, color = "black"), 
        axis.title = element_text(size = axis_title_size, face = "bold", color = "black")
        )

```

Calculate UMAP coordinates
```{r}
#load batch corrected, VST counts
norm_adjusted <- readRDS("~/Output_VSTcounts/batchcor_vstcounts_1298.rds") 

umap_out <- umap(t(norm_adjusted), random_state = 123, min_dist = 0.1, metric = 'cosine')
umap_2d <- data.frame(umap_out$layout) %>%
  tibble::rownames_to_column("samples")
umap_2d$samples <- gsub("^X", "", umap_2d$samples)
colnames(umap_2d) = c("coordinate_ID","UMAP1_2D", "UMAP2_2D")
umap_2d$UMAP2_2D <- -(umap_2d$UMAP2_2D)
write.csv(umap_2d, "./1298_umap2d_vst.csv")

# Merge with metadata
sample_info <-readxl::read_xlsx("~/DAM_data_clinical_patient_v40_20230808a.xlsx", skip = 5)
sample_info <- sample_info[-1, ]
sample_info_2 <- sample_info[match(umap_2d$coordinate_ID, sample_info$coordinate_ID), ] 

finaldf = left_join(umap_2d, sample_info_2,by="coordinate_ID" )
```

```{r}
setwd("./figures/")
```


DATASETS
```{r}
ggplot( finaldf %>% 
          arrange(factor(DATASET)),
        aes(
          x = UMAP1_2D, 
          y = UMAP2_2D,
          color = factor(DATASET),
        )
) + 
  geom_point(size=1.5) +
  theme_nolegend+
  ggtitle("Datasets") + 
  scale_color_manual(values = c("Baylor" = "blue",
                                "CAVATICA" = "grey",
                                "CHOP" = "forestgreen",
                                "Yale" = "chocolate",
                                "Heidelberg" = "purple",
                                "HKU/UCSF" = "deepskyblue",
                                "Palacky" = "black",
                                "UCSD" = "violet",
                                "UCSF(2018)" = "cyan",
                                "UCSF(2020)" = "gold", 
                                "UCSF(2022)" = "orange",
                                "UoT" = "red4",
                                "UW/FHCC" = "red"))
  

ggsave("DATASET_noleg.pdf", width = 7, height = 7)
```

CHR22 LOSS
```{r}
ggplot( finaldf %>% 
          arrange(factor(CHR22LOSS)),
        aes(
          x = UMAP1_2D, 
          y = UMAP2_2D,
          color = factor(CHR22LOSS),
        )
) + 
  geom_point(size=1.5) +
  theme_nolegend + 
  ggtitle("Chr 22 Loss") +
  scale_color_manual(values = c('yes' = "blue",
                                'no' = "orange",
                                "na" = "grey" 
  ))

ggsave("Chr22Loss_noleg.pdf", width = 7, height = 7)

```

NF2 Mutations
```{r}
ggplot( finaldf %>% 
          arrange(factor(NF2)),
        aes(
          x = UMAP1_2D, 
          y = UMAP2_2D,
          color = factor(NF2),
        )
) + 
  geom_point(size=1.5) +
  theme_nolegend + 
  ggtitle("NF2 Mutations") +
  scale_color_manual(values = c('yes' = "blue",
                                'no' = "orange",
                                "na" = "grey"))

ggsave("NF2mutations_noleg.pdf", width = 7, height = 7)
```

NF2 Fusions
```{r}
ggplot( finaldf %>% 
          arrange(factor(NF2_FUSION)),
        aes(
          x = UMAP1_2D, 
          y = UMAP2_2D,
          color = factor(NF2_FUSION),
        )
) + 
  geom_point(size=1.5) +
  theme_nolegend +
  ggtitle("NF2 Fusions") +
  scale_color_manual(values = c('yes' = "blue",
                                'no' = "orange",
                                "na" = "grey"))

ggsave("NF2Fusions_noleg.pdf", width = 7, height = 7)
```

WHO GRADE
```{r}
ggplot( finaldf %>% 
          arrange(factor(WHO_GRADE)),
        aes(
          x = UMAP1_2D, 
          y = UMAP2_2D,
          color = factor(WHO_GRADE),
        )
) + 
  geom_point(size=1.5) +
  theme_nolegend+
  ggtitle("WHO Grade") + 
  scale_color_manual(values = c(
    'WHO 1' = "orange",
    'WHO 2' = "forestgreen",
    'WHO 3' = "red",
    "na" = "grey"
  ))

setwd("./figures/")
ggsave("WHO_GRADE_vst.pdf", width = 7, height = 7 )
```

YAP1 Fusions
```{r}
MAML <- subset(finaldf, finaldf$YAP1_Fusion == "YAP1-MAML2")
LMO <- subset(finaldf, finaldf$YAP1_Fusion == "YAP1-LMO1")
ggplot( finaldf %>% 
          arrange(desc(factor(YAP1_Fusion))),
        aes(
          x = UMAP1_2D, 
          y = UMAP2_2D,
          color = factor(YAP1_Fusion),
        )
) + 
  geom_point(size=1.5) +
  geom_point(data=MAML,
             aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(YAP1_Fusion)))+
  geom_point(data=LMO,
             aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(YAP1_Fusion)))+
  theme_nolegend + 
  ggtitle("YAP1 Fusions") +
  scale_color_manual(values = c("YAP1-MAML2" = "red",
                                "YAP1-LMO1" = "blue",
                                "na" = "grey"))




ggsave("YAP1Fusions_noleg.pdf", width = 7, height = 7)


```

TRAF7 KLF4 AKT1 SMO mutations
```{r}
ggplot( finaldf %>% 
          arrange(factor(TRAF7_AKT1_KLF4_SMO)),
        aes(
          x = UMAP1_2D, 
          y = UMAP2_2D,
          color = factor(TRAF7_AKT1_KLF4_SMO),
        )
) + 
  geom_point(size=1.5) +
  theme_nolegend + 
  ggtitle("TRAF7 AKT1 KLF4 SMO") +
  scale_color_manual(values = c('yes' = "blue",
                                'no' = "orange",
                                "na" = "grey"))

ggsave("TKASmutations_noleg_vst.pdf", width = 7, height = 7)
```

Gender
```{r}
# Gender
MALE <- subset(finaldf, finaldf$GENDER == "male")
ggplot( finaldf %>% 
          arrange(desc(factor(GENDER))),
        aes(
          x = UMAP1_2D, 
          y = UMAP2_2D,
          color = factor(GENDER),
        )
) + 
  geom_point(size=1.5) +
  geom_point(data=MALE,
             aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(GENDER)))+
  theme_nolegend + 
  ggtitle("Gender") +
  scale_color_manual(values = c("male" = "blue",
                                "female" = "violet",
                                "na" = "grey"))
ggsave("Gender_noleg_vst.pdf", width = 7, height = 7)
```

Gender in 2 regions (Region 1 and rest of the UMAP) 
```{r}

Reg1 <- read.csv("./Region1_gender_Onc92VST.csv")
Reg1$Patient <- toupper(Reg1$Patient)

#rename deidentified samples
deID <- read.csv("~/Deidentification/SampleID_new.csv")
colnames(Reg1)[1] <- "sampleID_new"
Reg1_coID <- left_join(Reg1, deID, by="sampleID_new")

# Assign Region labels based on coordinate_ID
finaldf$Region[finaldf$coordinate_ID %in% Reg1_coID$coordinate_ID] <- "Region 1"
finaldf$Region[!(finaldf$coordinate_ID %in% Reg1_coID$coordinate_ID)] <- "All-Region1"

finaldf <- finaldf[!is.na(finaldf$Region), ]

#Group and summarize data
BG_tbl <- finaldf %>% group_by(Region, GENDER) %>% 
  summarise(total_count=n(),
            .groups = "drop") %>% 
  as.data.frame()

#Calculate counts for Region X and All-X
Region_X_count <- sum(BG_tbl[which(BG_tbl$Region == "Region X"), 3])
Region_AllX_count <- sum(BG_tbl[which(BG_tbl$Region == "All-X"), 3])

# Create table without rows 3 and 6 (remove na counts)
BG_tbl_2 <- BG_tbl[-c(3,6), ]
perc <- c(69,31,39,61) #manually calculated percentages
BG_tbl_2$percentage <- perc

#Create a bar plot (Supplemental Figure)
ggplot(BG_tbl_2, aes(x=factor(Region, level=c("Region 1", "All-Region1")), y=percentage))+
  geom_bar(aes(fill=GENDER), stat="identity", position= position_dodge(preserve = "single"), color="black") +
  theme_grid_legend +
  ggtitle("Gender")+
  scale_fill_manual(values = c('male' = "blue",
                               'female' = "violet",
                               "na"  = "grey"
  )) +
  xlab("Regions")

ggsave("gender_region1_noleg_vst.pdf", height = 7, width = 10)

```


Recurrent Tumor
```{r}
Rec <- subset(finaldf, finaldf$has_the_tumor_reccur == "yes")
nonRec <- subset(finaldf, finaldf$has_the_tumor_reccur == "no")
ggplot(finaldf,
       aes(
         x = UMAP1_2D, 
         y = UMAP2_2D,
         color = factor(has_the_tumor_reccur),
       )
) + 
  geom_point(size=1.5) +
  geom_point(data=nonRec,
             aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(has_the_tumor_reccur))) +
  geom_point(data=Rec,
           aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(has_the_tumor_reccur))) +
  theme_nolegend + 
  ggtitle("Recurrent Tumors")  +
  scale_color_manual(values = c(
    'yes' = "blue",
    'no' = "orange",
    "na" = "grey"
  ))

ggsave("Recurrent_Tumors_vst.pdf", width = 7, height = 7)
```

Age categories - umap 
```{r}
finaldf$AGE <- as.numeric(finaldf$AGE)
finaldf$AGE_CATEGORY <- cut(finaldf$AGE, breaks = c(-1, 0, 9, 18, 30, 60, 90, 98),
                            labels = c("1--9", "1--9", "10--18", "19--30","31--60", "61--90", "91--98"))

finaldf$AGE_CATEGORY <- as.character(finaldf$AGE_CATEGORY)
finaldf$AGE_CATEGORY[is.na(finaldf$AGE_CATEGORY)] <- "na"
finaldf$AGE_CATEGORY <- as.factor(finaldf$AGE_CATEGORY)

One <- subset(finaldf, finaldf$AGE_CATEGORY == "1--9")
Two <- subset(finaldf, finaldf$AGE_CATEGORY == "10--18")
Three <- subset(finaldf, finaldf$AGE_CATEGORY == "19--30")
Four <- subset(finaldf, finaldf$AGE_CATEGORY == "31--60")
Five <- subset(finaldf, finaldf$AGE_CATEGORY == "61--90")
Six <- subset(finaldf, finaldf$AGE_CATEGORY == "91--98")
Seven <- subset(finaldf, finaldf$AGE_CATEGORY == "na")

ggplot(finaldf %>% 
         arrange(factor(AGE_CATEGORY)), aes(x = UMAP1_2D, y = UMAP2_2D)) + 
  geom_point(size=1.5) +
  theme_nolegend +
  ggtitle("Age") + 
  geom_point(data = Seven, aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(AGE_CATEGORY))) +
  geom_point(data = Six, aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(AGE_CATEGORY))) +
  geom_point(data = Five, aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(AGE_CATEGORY))) +
  geom_point(data = Four, aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(AGE_CATEGORY))) +
  geom_point(data = Three, aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(AGE_CATEGORY))) +
  geom_point(data = Two, aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(AGE_CATEGORY))) +
  geom_point(data = One, aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(AGE_CATEGORY))) +
  scale_color_manual(values = c(
    '1--9' = " blue",
    '10--18' = "firebrick1 ",
    "19--30" = "sandybrown",
    "31--60" = " khaki3",
    "61--90" = " lavenderblush3", 
    "91--98" = "paleturquoise4 ", 
    "na"  = "grey39"
  ))
ggsave("age_categories_noleg_vst.pdf", width = 7, height = 7)

```

Bar graph showing Age Categories in 2 regions (Region 2 and rest of the UMAP) (Supplemental Figure)
```{r}

Reg1 <- read.csv("./Region2_age_Onc92VST.csv") #downloaded samples in Region 2 from Oncoscape 2D_VST plot
Reg1$Patient <- toupper(Reg1$Patient)

#rename deidentified samples
deID <- read.csv("~/Deidentification/SampleID_new.csv")
colnames(Reg1)[1] <- "sampleID_new"
Reg1_coID <- left_join(Reg1, deID, by="sampleID_new")

# Assign Region labels based on coordinate_ID
finaldf$Region[finaldf$coordinate_ID %in% Reg1_coID$coordinate_ID] <- "Region 2"
finaldf$Region[!(finaldf$coordinate_ID %in% Reg1_coID$coordinate_ID)] <- "All-Region2"

finaldf <- finaldf[!is.na(finaldf$Region), ]
finaldf$AGE <- as.numeric(finaldf$AGE)
finaldf$AGE_CATEGORY <- cut(finaldf$AGE, breaks = c(-1, 0, 9, 18, 30, 60, 90, 98),
                            labels = c("1--9", "1--9", "10--18", "19--30","31--60", "61--90", "91--98"))
finaldf$AGE_CATEGORY <- as.character(finaldf$AGE_CATEGORY)
finaldf$AGE_CATEGORY[is.na(finaldf$AGE_CATEGORY)] <- "na"
finaldf$AGE_CATEGORY <- as.factor(finaldf$AGE_CATEGORY)

#Group and summarize data
BG_tbl <- finaldf %>% group_by(Region, AGE_CATEGORY) %>% 
  summarise(total_count=n(),
            .groups = "drop") %>% 
  as.data.frame()

#Calculate counts in Region X and All-X
Region_X_count <- sum(BG_tbl[which(BG_tbl$Region == "Region 2"), 3])
Region_AllX_count <- sum(BG_tbl[which(BG_tbl$Region == "All-Region2"), 3])

#calculate percentages for each category
BG_tbl_2 <- BG_tbl[-c(7,14), ] #remove na
BG_tbl_3 <- BG_tbl_2 %>% filter(Region != "All-Region2") %>% mutate(Percentage = total_count/sum(total_count)*100)
BG_tbl_4 <- BG_tbl_2 %>% filter(Region != "Region 2") %>% mutate(Percentage = total_count/sum(total_count)*100)
BG_tbl_5 <- rbind(BG_tbl_3, BG_tbl_4)

#Generate bar plot
ggplot(BG_tbl_5, aes(x=factor(Region, level=c("Region 2", "All-Region2")), y=Percentage))+
  geom_bar(aes(fill=AGE_CATEGORY), stat="identity", position= position_dodge(preserve = "single"), color="black") +
  theme_grid_nolegend +
  ggtitle("AGE")+
  scale_fill_manual(values = c('1--9' = " blue",
                               '10--18' = "firebrick1 ",
                               "19--30" = "sandybrown",
                               "31--60" = " khaki3",
                               "61--90" = " lavenderblush3", 
                               "91--98" = "paleturquoise4 ", 
                               "na"  = "grey39"
  )) +
  xlab("Regions")

ggsave("Age_Region2_vst_noleg.pdf", height = 7, width = 10)
```


FFPE and Frozen
```{r}
ggplot( finaldf %>% 
  arrange(desc(factor(FFPE_or_Frozen))),
aes(
  x = UMAP1_2D, 
  y = UMAP2_2D,
  color = factor(FFPE_or_Frozen),
)
) + 
  geom_point(size=1.5) +
  theme_nolegend + 
  ggtitle("FFPE and frozen samples") +
  scale_color_manual(values = c("FFPE" = "blue",
                                "frozen" = "grey80",
                                "na" = "grey"))
setwd("./figures/")
#ggsave("FFPE_frozen_leg.pdf", width = 10, height = 7)
ggsave("FFPE_frozen_noleg.pdf", width = 7, height = 7)
```

NF2 Mutations + NF2 Fusions + Chr22 Loss
```{r}
ggplot( finaldf %>% 
          arrange(factor(NF2_ISSUE)),
        aes(
          x = UMAP1_2D, 
          y = UMAP2_2D,
          color = factor(NF2_ISSUE),
        )
) + 
  geom_point(size=1.5) +
  theme_nolegend + 
  ggtitle("NF2 Mut + NF2 Fusions + Chr 22 Loss ")  +
  scale_color_manual(values = c('yes' = "blue",
                                'no' = "orange",
                                "na" = "grey"))

#ggsave("FFPE_frozen_leg.pdf", width = 10, height = 7)
ggsave("NF2Mut_Fus_chr22loss_noleg.pdf", width = 7, height = 7)
```

Tumors with more than one of TRAF7, KLF4, AKT1, SMO mutations
```{r}
finaldf$nonNF2[finaldf$TRAF7 == "na"] <- "na"
finaldf$nonNF2[finaldf$KLF4 == "na"] <- "na"
finaldf$nonNF2[finaldf$SMO == "na"] <- "na"
finaldf$nonNF2[finaldf$AKT1 == "na"] <- "na"
finaldf$nonNF2[finaldf$TRAF7 == "no"] <- "na"
finaldf$nonNF2[finaldf$KLF4 == "no"] <- "na"
finaldf$nonNF2[finaldf$SMO == "no"] <- "na"
finaldf$nonNF2[finaldf$AKT1 == "no"] <- "na"
finaldf$nonNF2[finaldf$TRAF7 == "yes"] <- "TRAF7"
finaldf$nonNF2[finaldf$KLF4 == "yes"] <- "KLF4"
finaldf$nonNF2[finaldf$SMO == "yes"] <- "SMO"
finaldf$nonNF2[finaldf$AKT1 == "yes"] <- "AKT1"
finaldf$nonNF2[finaldf$TRAF7 == "yes" & finaldf$KLF4=="yes"] <- "TRAF7, KLF4"
finaldf$nonNF2[finaldf$TRAF7 == "yes" & finaldf$AKT1=="yes"] <- "TRAF7, AKT1"
finaldf$nonNF2[finaldf$TRAF7 == "yes" & finaldf$SMO=="yes"] <- "TRAF7, SMO"
finaldf$nonNF2[finaldf$TRAF7 == "yes" & finaldf$KLF4=="yes" & finaldf$SMO=="yes"] <- "TRAF7, KLF4, SMO"
finaldf$nonNF2[finaldf$TRAF7 == "yes" & finaldf$NF2=="yes"] <- "TRAF7, NF2"

ggplot( finaldf %>% 
          arrange(factor(nonNF2)),
        aes(
          x = UMAP1_2D, 
          y = UMAP2_2D,
          color = factor(nonNF2),
        )
) + 
  geom_point(size=1.5) +
  theme_nolegend + 
  ggtitle("TRAF7 AKT1 KLF4 SMO") +
  scale_color_manual(values = c("TRAF7" = "blue",
                                "KLF4" = "red",
                                "AKT1" = "green",
                                "SMO" = "brown",
                                "TRAF7, KLF4" = "purple",
                                "TRAF7, AKT1" = "deepskyblue",
                                "TRAF7, SMO" = "black",
                                "TRAF7, KLF4, SMO" = "violet",
                                "TRAF7, NF2" = "cyan",
                                "no" = "gold",
                                "na" = "grey"))

ggsave("TKAS_multiplemutations_vst_noleg.pdf", width = 7, height = 7)
```

KLF4 and AKT1 mutations
```{r}
finaldf$KLF4AKT1[finaldf$TRAF7 == "na"] <- "na"
finaldf$KLF4AKT1[finaldf$KLF4 == "na"] <- "na"
finaldf$KLF4AKT1[finaldf$SMO == "na"] <- "na"
finaldf$KLF4AKT1[finaldf$AKT1 == "na"] <- "na"
finaldf$KLF4AKT1[finaldf$TRAF7 == "no"] <- "na"
finaldf$KLF4AKT1[finaldf$KLF4 == "no"] <- "na"
finaldf$KLF4AKT1[finaldf$SMO == "no"] <- "na"
finaldf$KLF4AKT1[finaldf$AKT1 == "no"] <- "na"
finaldf$KLF4AKT1[finaldf$KLF4 == "yes"] <- "KLF4"
finaldf$KLF4AKT1[finaldf$AKT1 == "yes"] <- "AKT1"
KLF4 <- subset(finaldf, finaldf$KLF4AKT1 == "KLF4")
AKT1 <- subset(finaldf, finaldf$KLF4AKT1 == "AKT1")

ggplot( finaldf %>% 
          arrange(factor(KLF4AKT1)),
        aes(
          x = UMAP1_2D, 
          y = UMAP2_2D,
          color = factor(KLF4AKT1),
        )
) + 
  geom_point(size=1.5) +
  geom_point(data=KLF4, aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(KLF4AKT1))) +
  geom_point(data=AKT1, aes(x=UMAP1_2D, y=UMAP2_2D, color=factor(KLF4AKT1))) +
  theme_nolegend + 
  ggtitle("KLF4 and AKT1 mutations") +
  scale_color_manual(values = c(
    "KLF4" = "red",
    "AKT1" = "blue",
    "na" = "grey"))

ggsave("KLF4_AKT1_vst_noleg.pdf", width = 7, height = 7)
```

## Months to recurrence
```{r}
sample_1298 <- subset(sample_info, sample_info$coordinate_ID %in% colnames(vstcounts))
sample_timetorec <- subset(sample_1298, !(sample_1298$has_the_tumor_progressed == "yes")) %>% subset(., !(.$time_for_KM_final == "na"))
#sample_timetorec <- sample_1298
sample_timetorec$time_for_KM_final <- as.numeric(as.character(sample_timetorec$time_for_KM_final))

sample_timetorec$time_for_KM_final_category <- cut(sample_timetorec$time_for_KM_final, breaks = c(-1, 0, 24, 60, 120, 315),
                                                   labels = c("0-24", "0-24", "25-60", "61-120", "120<"))

sample_info_sub <- sample_timetorec[, c("coordinate_ID", "time_for_KM_final_category", "has_the_tumor_reccur", "time_for_KM_final")]
sample_info_sub_2 <- subset(sample_info_sub, sample_info_sub$has_the_tumor_reccur == "yes")

fianldf <- left_join(finaldf, sample_info_sub_2, by="coordinate_ID")

fianldf$time_for_KM_final_category <- as.character(fianldf$time_for_KM_final_category)
fianldf$time_for_KM_final_category[is.na(fianldf$time_for_KM_final_category)] <- "na"
fianldf$time_for_KM_final_category <- as.factor(fianldf$time_for_KM_final_category)

One <- subset(fianldf, fianldf$time_for_KM_final_category == "0-24" & fianldf$has_the_tumor_reccur == "yes")
Two <- subset(fianldf, fianldf$time_for_KM_final_category == "25-60" & fianldf$has_the_tumor_reccur == "yes")
Three <- subset(fianldf, fianldf$time_for_KM_final_category == "61-120" & fianldf$has_the_tumor_reccur == "yes")
Four <- subset(fianldf, fianldf$time_for_KM_final_category == "120<" & fianldf$has_the_tumor_reccur == "yes")
Five <- subset(fianldf, fianldf$time_for_KM_final_category == "na")
ggplot( fianldf %>% 
          arrange(desc(factor(time_for_KM_final_category))),
        aes(
          x = UMAP1_2D, 
          y = UMAP2_2D,
          color = factor(time_for_KM_final_category),
        )
) + 
  geom_point( size=1.5) +
  theme_nolegend +
  ggtitle("MONTHS TO RECURRENCE") +
  scale_color_manual(values = c( 
    "0-24" = "red",
    "25-60" = "gold",
    "61-120" = "green4",
    "120<" = "royalblue1", 
    "na" = "grey"))

ggsave("MonthsToRec_noleg.pdf", width = 7, height = 7)

```

