----------------------
title: "Differential Gene Expression analysis to determine upregulated genes in each cluster"
authors: "Nayanga Thirimanne"
----------------------

```{r}
library(edgeR)
```

load cluster IDs from dbscan analysis
```{r}
clusterID <- read_csv("./DBSCAN/dbscan2D_un37_vst.csv")
cluster_1 <- clusterID[clusterID$cluster == "1",]
cluster_2 <- clusterID[clusterID$cluster == "2",]
cluster_3 <- clusterID[clusterID$cluster == "3",]
cluster_4 <- clusterID[clusterID$cluster == "4",]
cluster_5 <- clusterID[clusterID$cluster == "5",]
cluster_6 <- clusterID[clusterID$cluster == "6",]
cluster_7 <- clusterID[clusterID$cluster == "7",]
```

load batch corrected raw counts
```{r}
rawcounts1 <- readRDS("~/1298_combatseq_rawcounts.rds")
```

remove unclustered samples
```{r}
unclust <- clusterID[clusterID$cluster == "0",]
clusterID <- subset(clusterID, !(clusterID$coordinate_ID %in% unclust$coordinate_ID))
rawcounts1 <- rawcounts1[, !(colnames(rawcounts1) %in% unclust$coordinate_ID)]
```

DEG between Cluster A and rest of the samples in the UMAP
```{r}
#generate sample dataframe
sample <- colnames(rawcounts1)
Adf <- data.frame(sample)
Adf$group[Adf$sample %in% cluster_1$coordinate_ID] <- 'A'
Adf$group[!(Adf$sample %in% cluster_1$coordinate_ID)] <- 'All'
#EdgeR DE analysis: GLM approach (quasi-likelihood F-tests)
#set group
group = factor(Adf$group, levels = c("A","All"))
cols_keep <- Adf$sample
rawcounts <- rawcounts1[, colnames(rawcounts1) %in% cols_keep]
y <- DGEList(counts = rawcounts, genes = rownames(rawcounts), samples = colnames(rawcounts), group = group)
y <- calcNormFactors(y)
#set design
design <- model.matrix(~0+group)
rownames(design) <- colnames(y)
#estimate the NB disperson for the dataset
y <- estimateDisp(y, design, robust = TRUE)
y$common.dispersion
y$tagwise.dispersion
plotBCV(y) 
fit <- glmQLFit(y, design)
my.contrasts1 <- makeContrasts(AvsAll=groupA-groupAll, levels=design)
qlf <- glmQLFTest(fit, contrast = my.contrasts1) 
topTags(qlf)
W <- as.data.frame(topTags(qlf), n=nrow(y))
out <- topTags(qlf, n=Inf)
AvsAll <- as.data.frame(out)
write.csv(AvsAll, "./DE_dbs_clusterAvsAll_vst_unfilt.csv")

```

DEG between cluster B and all other samples in the UMAP
```{r}
#generate sample dataframe
sample <- colnames(rawcounts1)
Adf <- data.frame(sample)
Adf$group[Adf$sample %in% cluster_2$coordinate_ID] <- 'B'
Adf$group[!(Adf$sample %in% cluster_2$coordinate_ID)] <- 'All'
#EdgeR DE analysis: GLM approach (quasi-likelihood F-tests)
#set group
group = factor(Adf$group, levels = c("B","All"))
cols_keep <- Adf$sample
rawcounts <- rawcounts1[, colnames(rawcounts1) %in% cols_keep]
y <- DGEList(counts = rawcounts, genes = rownames(rawcounts), samples = colnames(rawcounts), group = group)
y <- calcNormFactors(y)
#set design
design <- model.matrix(~0+group)
rownames(design) <- colnames(y)
#estimate the NB disperson for the dataset
y <- estimateDisp(y, design, robust = TRUE)
y$common.dispersion
y$tagwise.dispersion
plotBCV(y) 
fit <- glmQLFit(y, design)
my.contrasts1 <- makeContrasts(BvsAll=groupB-groupAll, levels=design)
qlf <- glmQLFTest(fit, contrast = my.contrasts1) 
topTags(qlf)
W <- as.data.frame(topTags(qlf), n=nrow(y))
out <- topTags(qlf, n=Inf)
BvsAll <- as.data.frame(out)
write.csv(BvsAll, "./DE_dbs_clusterBvsAll_vst_unfilt.csv")

```


DEG between cluster E and all other samples in the UMAP
```{r}
#generate sample dataframe
sample <- colnames(rawcounts1)
Adf <- data.frame(sample)
Adf$group[Adf$sample %in% cluster_5$coordinate_ID] <- 'E'
Adf$group[!(Adf$sample %in% cluster_5$coordinate_ID)] <- 'All'
#EdgeR DE analysis: GLM approach (quasi-likelihood F-tests)
#set group
group = factor(Adf$group, levels = c("E","All"))
cols_keep <- Adf$sample
rawcounts <- rawcounts1[, colnames(rawcounts1) %in% cols_keep]
y <- DGEList(counts = rawcounts, genes = rownames(rawcounts), samples = colnames(rawcounts), group = group)
y <- calcNormFactors(y)
#set design
design <- model.matrix(~0+group)
rownames(design) <- colnames(y)
#estimate the NB disperson for the dataset
y <- estimateDisp(y, design, robust = TRUE)
y$common.dispersion
y$tagwise.dispersion
plotBCV(y) 
fit <- glmQLFit(y, design)
my.contrasts1 <- makeContrasts(EvsAll=groupE-groupAll, levels=design)
qlf <- glmQLFTest(fit, contrast = my.contrasts1) 
topTags(qlf)
W <- as.data.frame(topTags(qlf), n=nrow(y))
out <- topTags(qlf, n=Inf)
EvsAll <- as.data.frame(out)
write.csv(EvsAll, "./DE_dbs_clusterEvsAll_vst_unfilt.csv")

```
```


Above DEG analysis was performed on all the other clusters (e.g. Cluster B vs "All other samples")


Pathway enrichment analysis
```{r}
DE <- AvsAll
Upregulated_genes <- DE[DE$logFC <= 0.6 & DE$FDR <= 0.05, ]
#Upregulated genes in each cluster were run on Enrichr.
#Gene Ontology Biological Processes (GOBP) terms enriched in each cluster were downloaded from Enrichr
#Top 15 GOBP terms were plotted as below

plot_title_size = 12
legend_pt_size = 6
axis_text_size = 100
axis_title_size = 15
legend_text_size = 10
spacing = 0.5
chosen_margin = c(0.5,1,0.5,1) #top,right,bottom,left

theme_grid_nolegend <- theme_bw()+
  theme(plot.title = element_text(hjust=0, vjust=0, lineheight=.8, face="bold", size=plot_title_size),
        plot.margin = unit(chosen_margin,"cm"),
        legend.text = element_text(size=legend_text_size),
        legend.key.height = unit(spacing,"cm"),
        legend.position = "none",
        legend.justification = "left",
        legend.title = element_blank())

theme_grid_legend <- theme_bw()+
  theme(plot.title = element_text(hjust=0, vjust=0, lineheight=.8, face="bold", size=plot_title_size),
        plot.margin = unit(chosen_margin,"cm"),
        legend.text = element_text(size=legend_text_size),
        legend.key.height = unit(spacing,"cm"),
        legend.position = "right",
        legend.justification = "left",
        legend.title = element_text())

En_A <- read.csv("~/Analysis27_DE_1298/Enrichr_output/Enrichr_GO_BP_ClA_edited.csv")
En_A_2 <- top_n(En_A, n = 15, -p.adjust)
En_A_2$Term <- str_replace(En_A_2$Term, " \\s*\\([^\\)]+\\)", "")
En_A_2$Term <- factor(En_A_2$Term, levels = En_A_2$Term)
ggplot(En_A_2, aes(x=-log10(p.adjust), y= reorder(Term, desc(Term)), size=GeneRatio, color=-log10(p.adjust)))+
  geom_point(stat = "identity") +
  scale_color_gradient(low="blue",high="red") +
  ylab("") +
  theme_grid_legend +
  xlab("-log10(p.adjust)")+
  ggtitle("Cluster A")
ggsave("ClusterA_leg.pdf", width = 8, height = 3.5)


```

